---
title: "Experiment 2: Adult and Child directed LLM promt comparison"
output: html_notebook
---

### Load libraries
```{r}
library(ggplot2)
library(reshape2)
library(lme4)
# Please note, we also use functions from cowplot and plyr, thus please install these packages to run this notebook
```

### Load DeveL data and converge long format

DeveL including
- SUBTLEX-Ger
- LLM based Frequencies

Including a group variable that categorizes Grade 1-4 into *Kids*, Grade 6 into *Adolescent* and young as well as old adults into *adults*. 

```{r}
load("devel-LLM_Experiment_2.RData",verbose = T)

#convert to long format plus inclusion of the group variable
long_merged_df = na.omit(data.frame(rt = log(c(results_merged$rt.g1.m
                                               ,results_merged$rt.g2.m
                                               ,results_merged$rt.g3.m
                                               ,results_merged$rt.g4.m
                                               ,results_merged$rt.g6.m
                                               ,results_merged$rt.ya.m
                                               ,results_merged$rt.oa.m)
                                             )
                            , group = as.factor(c(rep("Kids",length(c(results_merged$rt.g1.m
                                                                      ,results_merged$rt.g2.m
                                                                      ,results_merged$rt.g3.m
                                                                      ,results_merged$rt.g4.m)
                                                                    )
                                                      )
                                                  ,rep("Adolescent",length(results_merged$rt.g6.m)
                                                       )
                                                  ,rep("Adults",length(c(results_merged$rt.ya.m
                                                                         ,results_merged$rt.oa.m)
                                                                       )
                                                       )
                                                  )
                                                )
                            , age = c(rep(1,length(results_merged$rt.g1.m))
                                       ,rep(2,length(results_merged$rt.g2.m))
                                       ,rep(3,length(results_merged$rt.g3.m))
                                       ,rep(4,length(results_merged$rt.g4.m))
                                       ,rep(5,length(results_merged$rt.g6.m))
                                       ,rep(6,length(results_merged$rt.ya.m))
                                       ,rep(7,length(results_merged$rt.oa.m))
                                       )
                            , child.old20 = rep(results_merged$child.old20, 7)
                            , aoa = rep(results_merged$aoa, 7)
                            , letter.cnt = rep(results_merged$letter.cnt, 7)
                            , llm_freq_log = rep(results_merged$log_frequency_per_million_b_LLM_exp1_j, 7)
                            , log_frequency_per_million_hi_kids = rep(results_merged$log_frequency_per_million_j_kids_hi_t_j, 7)
                            , log_frequency_per_million_lo_kids = rep(results_merged$log_frequency_per_million_j_kids_lo_t_j, 7)
                            , log_frequency_per_million_hi_adults = rep(results_merged$log_frequency_per_million_j_adults_hi_t_j, 7)
                            , log_frequency_per_million_lo_adults = rep(results_merged$log_frequency_per_million_j_adults_lo_t_j, 7)
                            , dwds_freq_log = rep(results_merged$log_frequency_per_million_b_dwds_j, 7)
                            , childlex_freq_log = rep(results_merged$log_frequency_per_million_b_childlex_j, 7)
                            , word = rep(results_merged$word, 7)
                      ))
summary(long_merged_df)
```

### Model estimation
Linear mixed models to estimate the change in model fit (AIC difference to a null model without word frequency) and estimated effect sizes for *Children* (Kids) and *Adults*.

#### Adult Models
```{r}
#Null Model Adults
m0a = lmer(rt~child.old20+aoa+letter.cnt+age+(1|word)
           ,data=subset(long_merged_df,long_merged_df$group=="Adults")
           )
#DWDS Model Adults
mSUBa = lmer(rt~child.old20+aoa+letter.cnt+age+dwds_freq_log+(1|word)
             ,data=subset(long_merged_df,long_merged_df$group=="Adults")
             )
#ChildLex Model Adults
mCHLEXa = lmer(rt~child.old20+aoa+letter.cnt+age+childlex_freq_log+(1|word)
               ,data=subset(long_merged_df,long_merged_df$group=="Adults")
               )
#LLM Freq. Exp. 1 Model Adults
moLLMa = lmer(rt~child.old20+aoa+letter.cnt+age+llm_freq_log+(1|word)
              ,data=subset(long_merged_df,long_merged_df$group=="Adults")
              )
#LLM Freq. Exp. 2 Model Adults: Low Temp. Adult directed
mfltaa = lmer(rt~child.old20+aoa+letter.cnt+age+log_frequency_per_million_lo_adults+(1|word)
              ,data=subset(long_merged_df,long_merged_df$group=="Adults")
              )
#LLM Freq. Exp. 2 Model Adults: High Temp. Adult directed
mfhtaa = lmer(rt~child.old20+aoa+letter.cnt+age+log_frequency_per_million_hi_adults+(1|word)
              ,data=subset(long_merged_df,long_merged_df$group=="Adults")
              )
#LLM Freq. Exp. 2 Model Adults: Low Temp. Child directed
mfltka = lmer(rt~child.old20+aoa+letter.cnt+age+log_frequency_per_million_lo_kids+(1|word)
              ,data=subset(long_merged_df,long_merged_df$group=="Adults")
              )
#LLM Freq. Exp. 2 Model Adults: High Temp. Child directed
mfhtka = lmer(rt~child.old20+aoa+letter.cnt+age+log_frequency_per_million_hi_kids+(1|word)
              ,data=subset(long_merged_df,long_merged_df$group=="Adults")
              )
```

#### Children Models
```{r}
#Null Model Children
m0k = lmer(rt~child.old20+aoa+letter.cnt+age+(1|word)
           ,data=subset(long_merged_df,long_merged_df$group=="Kids")
           )
#DWDS Model Children
mSUBk = lmer(rt~child.old20+aoa+letter.cnt+age+dwds_freq_log+(1|word)
             ,data=subset(long_merged_df,long_merged_df$group=="Kids")
             )
#ChildLex Model Children
mCHLEXk = lmer(rt~child.old20+aoa+letter.cnt+age+childlex_freq_log+(1|word)
               ,data=subset(long_merged_df,long_merged_df$group=="Kids")
               )
#LLM Freq. Exp. 1 Model Children
moLLMk = lmer(rt~child.old20+aoa+letter.cnt+age+llm_freq_log+(1|word)
              ,data=subset(long_merged_df,long_merged_df$group=="Kids")
              )
#LLM Freq. Exp. 2 Model Children: Low Temp. Adult directed
mfltak = lmer(rt~child.old20+aoa+letter.cnt+age+log_frequency_per_million_lo_adults+(1|word)
   ,data=subset(long_merged_df,long_merged_df$group=="Kids")
   )
#LLM Freq. Exp. 2 Model Children: High Temp. Adult directed
mfhtak = lmer(rt~child.old20+aoa+letter.cnt+age+log_frequency_per_million_hi_adults+(1|word)
   ,data=subset(long_merged_df,long_merged_df$group=="Kids")
   )
#LLM Freq. Exp. 2 Model Children: Low Temp. Children directed
mfltkk = lmer(rt~child.old20+aoa+letter.cnt+age+log_frequency_per_million_lo_kids+(1|word)
   ,data=subset(long_merged_df,long_merged_df$group=="Kids")
   )
#LLM Freq. Exp. 2 Model Children: High Temp. Children directed
mfhtkk = lmer(rt~child.old20+aoa+letter.cnt+age+log_frequency_per_million_hi_kids+(1|word)
   ,data=subset(long_merged_df,long_merged_df$group=="Kids") 
   )
```

#### Preprocessing for plotting
Calculation of Model fit (Delta AIC) and effect sizes (including the 95% confidence interval) for the frequency effect measured with alternative frequency measures. 
```{r}
aic_df = data.frame(deltaAIC = c(AIC(m0k) - AIC(mfltak)
                        ,AIC(m0k) - AIC(mfhtak)
                        ,AIC(m0k) - AIC(mfltkk)
                        ,AIC(m0k) - AIC(mfhtkk)
                        ,AIC(m0a) - AIC(mfltaa)
                        ,AIC(m0a) - AIC(mfhtaa)
                        ,AIC(m0a) - AIC(mfltka)
                        ,AIC(m0a) - AIC(mfhtka)
                        )
           ,efs = c(fixef(mfltak)["log_frequency_per_million_lo_adults"]
                        ,fixef(mfhtak)["log_frequency_per_million_hi_adults"]
                        ,fixef(mfltkk)["log_frequency_per_million_lo_kids"]
                        ,fixef(mfhtkk)["log_frequency_per_million_hi_kids"]
                        ,fixef(mfltaa)["log_frequency_per_million_lo_adults"]
                        ,fixef(mfhtaa)["log_frequency_per_million_hi_adults"]
                        ,fixef(mfltka)["log_frequency_per_million_lo_kids"]
                        ,fixef(mfhtka)["log_frequency_per_million_hi_kids"]
                        )
           , ci_min = c(confint(mfltak, level = 0.95)["log_frequency_per_million_lo_adults",1]
                        ,confint(mfhtak, level = 0.95)["log_frequency_per_million_hi_adults",1]
                        ,confint(mfltkk, level = 0.95)["log_frequency_per_million_lo_kids",1]
                        ,confint(mfhtkk, level = 0.95)["log_frequency_per_million_hi_kids",1]
                        ,confint(mfltaa, level = 0.95)["log_frequency_per_million_lo_adults",1]
                        ,confint(mfhtaa, level = 0.95)["log_frequency_per_million_hi_adults",1]
                        ,confint(mfltka, level = 0.95)["log_frequency_per_million_lo_kids",1]
                        ,confint(mfhtka, level = 0.95)["log_frequency_per_million_hi_kids",1])
  , ci_max = c(confint(mfltak, level = 0.95)["log_frequency_per_million_lo_adults",2]
               ,confint(mfhtak, level = 0.95)["log_frequency_per_million_hi_adults",2]
               ,confint(mfltkk, level = 0.95)["log_frequency_per_million_lo_kids",2]
               ,confint(mfhtkk, level = 0.95)["log_frequency_per_million_hi_kids",2]
               ,confint(mfltaa, level = 0.95)["log_frequency_per_million_lo_adults",2]
               ,confint(mfhtaa, level = 0.95)["log_frequency_per_million_hi_adults",2]
               ,confint(mfltka, level = 0.95)["log_frequency_per_million_lo_kids",2]
               ,confint(mfhtka, level = 0.95)["log_frequency_per_million_hi_kids",2])
           ,Group = as.factor(c(rep("Children RT",4),rep("Adults RT",4)))
           ,Frequency = as.factor(c(rep("Adults",2),rep("Children",2),rep("Adults",2),rep("Children",2)))
           ,Temperature = as.factor(rep(c("Low","High"),4))
)


aic_df_kids_bl_comp_alt = data.frame(deltaAIC = c(
                        AIC(m0k) - AIC(mfhtkk)
                        ,AIC(m0k) - AIC(moLLMk)
                        ,AIC(m0k) - AIC(mCHLEXk)
                        ,AIC(m0k) - AIC(mfhtak)
                        ,AIC(m0k) - AIC(mSUBk)
                        )
                        ,efs = c(fixef(mfhtkk)["log_frequency_per_million_hi_kids"]
                        ,fixef(moLLMk)["llm_freq_log"]
                        ,fixef(mCHLEXk)["childlex_freq_log"]
                        ,fixef(mfhtak)["log_frequency_per_million_hi_adults"]
                        ,fixef(mSUBk)["dwds_freq_log"]
                        )
           , ci_min = c(confint(mfhtkk, level = 0.95)["log_frequency_per_million_hi_kids",1]
                        ,confint(moLLMk, level = 0.95)["llm_freq_log",1]
                        ,confint(mCHLEXk, level = 0.95)["childlex_freq_log",1]
                        ,confint(mfhtak, level = 0.95)["log_frequency_per_million_hi_adults",1]
                        ,confint(mSUBk, level = 0.95)["dwds_freq_log",1]
                        )
  , ci_max = c(confint(mfhtkk, level = 0.95)["log_frequency_per_million_hi_kids",2]
               ,confint(moLLMk, level = 0.95)["llm_freq_log",2]
               ,confint(mCHLEXk, level = 0.95)["childlex_freq_log",2]
               ,confint(mfhtak, level = 0.95)["log_frequency_per_million_hi_adults",2]
               ,confint(mSUBk, level = 0.95)["dwds_freq_log",2]
               )
           ,Group = as.factor(c(rep("Children RT",5)))
           ,Frequency = as.factor(c("Exp. 2: LLM Hi. T. Children","Exp. 1: LLM Children","childLex","Exp. 2: LLM Hi. T. Adults","DWDS"))
)
aic_df_adults_bl_comp_alt = data.frame(deltaAIC = c(
                        AIC(m0a) - AIC(mfhtaa)
                        ,AIC(m0a) - AIC(moLLMa)
                        ,AIC(m0a) - AIC(mCHLEXa)
                        ,AIC(m0a) - AIC(mfhtka)
                        ,AIC(m0a) - AIC(mSUBa)
                        )
                        ,efs = c(fixef(mfhtaa)["log_frequency_per_million_hi_adults"]
                        ,fixef(moLLMa)["llm_freq_log"]
                        ,fixef(mCHLEXa)["childlex_freq_log"]
                        ,fixef(mfhtka)["log_frequency_per_million_hi_kids"]
                        ,fixef(mSUBa)["dwds_freq_log"]
                        )
           , ci_min = c(confint(mfhtaa, level = 0.95)["log_frequency_per_million_hi_adults",1]
                        ,confint(moLLMa, level = 0.95)["llm_freq_log",1]
                        ,confint(mCHLEXa, level = 0.95)["childlex_freq_log",1]
                        ,confint(mfhtka, level = 0.95)["log_frequency_per_million_hi_kids",1]
                        ,confint(mSUBa, level = 0.95)["dwds_freq_log",1]
                        )
  , ci_max = c(confint(mfhtaa, level = 0.95)["log_frequency_per_million_hi_adults",2]
               ,confint(moLLMa, level = 0.95)["llm_freq_log",2]
               ,confint(mCHLEXa, level = 0.95)["childlex_freq_log",2]
               ,confint(mfhtka, level = 0.95)["log_frequency_per_million_hi_kids",2]
               ,confint(mSUBa, level = 0.95)["dwds_freq_log",2]
               )
           ,Group = as.factor(c(rep("Adults RT",5)))
           ,Frequency = as.factor(c("Exp. 2: LLM Hi. T. Adults","Exp. 1: LLM Children","childLex","Exp. 2: LLM Hi. T. Children","DWDS"))
)
```
#### Code for Fig. 8

```{r}
al=.9
comp_pl_df = rbind(aic_df_adults_bl_comp_alt,aic_df_kids_bl_comp_alt)
comp_pl_df = subset(comp_pl_df,comp_pl_df$Frequency!="Exp. 2: LLM Hi. T. Adults")
comp_pl_df = subset(comp_pl_df,comp_pl_df$Frequency!="Exp. 2: LLM Hi. T. Children")
comp_pl_df$pl_alpha = c(rep(al,2),rep(1,2),rep(al,2))


aic_df$Frequency_efs = paste("       ",aic_df$Frequency,aic_df$Temperature,"T.")
aic_df$pl_alpha = c(rep(al,3),1,al,1,rep(al,2))


aic_diff_exp2_ = ggplot(aic_df, aes(y= deltaAIC, x=Frequency_efs , fill=Temperature,alpha = pl_alpha))+
  geom_bar(stat = "identity",
           position = "dodge")+
  theme_minimal()+
  ylab("Model fit increase\n[AIC relative to H0]")+
  xlab("")+
  scale_color_manual(values = c("black","black"))+
  scale_alpha_manual(values = c(1,.7))+
  scale_fill_manual(values = c("#FF5580","#03045e"))+  facet_wrap(~Group)+
  geom_hline(yintercept = 3,color="white",lty=2)+
  ylim(0,365)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(legend.position = c(0, 1), 
        legend.justification = c(0, 1),
    legend.box.background = element_rect(fill = "white", 
                                         color = "white", 
                                         size = 1))+
  scale_alpha_continuous(range = c(0.3, 1)) +
  guides(alpha = "none")

all_ef_sizes_ = ggplot(aic_df, aes(x = Frequency_efs, y = efs, color = Temperature,alpha = pl_alpha)) +
  geom_errorbar(aes(ymin = ci_min, ymax = ci_max), width = 0.2) +
  geom_point() + 
  labs(x = "", y = "Frequency effect size\n[log. ms]") +
  theme_minimal() +ylim(-.026,0)+
  scale_color_manual(values = c("#FF5580","#03045e")) +  facet_wrap(~Group)+ theme(legend.position="none")+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  geom_hline(yintercept = 0,color="white",lty=2)+
  scale_alpha_continuous(range = c(0.3, 1)) +
  guides(alpha = "none")

aic_diff_exp2_kids_adults_ = ggplot(comp_pl_df, aes(y= deltaAIC, x=Frequency, fill=Frequency,alpha = pl_alpha))+
  geom_bar(stat = "identity",
           position = "dodge")+
  theme_minimal()+
  ylab("")+
  xlab("")+
  scale_color_manual(values = c("black","black"))+
  scale_alpha_manual(values = c(1,.7))+
  scale_fill_manual(values = c("#e63946","#2a9d8f","#0077b6"))+ 
  facet_wrap(~Group)+
  geom_hline(yintercept = 3,color="white",lty=2)+ 
  theme(legend.position="none")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+ylim(0,365)+
  scale_alpha_continuous(range = c(0.3, 1)) +
  guides(alpha = "none")

all_ef_sizes_kids_adults_ = ggplot(comp_pl_df, aes(x = Frequency, y = efs, color = Frequency,alpha = pl_alpha)) +
  geom_errorbar(aes(ymin = ci_min, ymax = ci_max), width = 0.2) +
  geom_point() + 
  labs(x = "", y = " ") +
  theme_minimal() +ylim(-.026,0)+
  scale_color_manual(values = c("#e63946","#2a9d8f","#0077b6")) +  facet_wrap(~Group)+ theme(legend.position="none")+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  geom_hline(yintercept = 0,color="white",lty=2)+
  scale_alpha_continuous(range = c(0.3, 1)) +
  guides(alpha = "none")

cowplot::plot_grid(aic_diff_exp2_,aic_diff_exp2_kids_adults_,all_ef_sizes_,all_ef_sizes_kids_adults_,nrow=2,labels = "AUTO",rel_widths = c(1,.75))
```